\Large \bf DRAFT: Direct-style process creation on Linux

### 2021-04-08 research/paper/paper.tex
paper

\begin{document}
\newcommand{\twosigma}{[tech company]} 
\date{}
\title{\Large \bf DRAFT: Direct-style process creation on Linux}
% \author{
% {\rm Spencer Baugh}\\
% Two Sigma
% }
\maketitle
\begin{abstract}
Traditional process creation interfaces,
such as fork and spawn,
are complex to use, limited in power, and difficult to abstract over.
We develop a new process creation interface for Linux
which allows a program to create a child process in an non-running state
and initialize the new process by operating on it from the outside.
This method of process creation results in more comprehensible programs, 
has better error handling,
is comparably efficient,
and is more amenable to abstraction.
Our implementation is immediately deployable without kernel modifications on any recent Linux kernel version.
\end{abstract}

@@

We evaluate the degree to which we meet this goal by answering the following questions:
\begin{compactitem}
\item
  Does direct-style process creation on Linux allow for simpler programs than the alternatives?
  We compare implementations of similar functionality using both direct-style and fork-style in \ref{ease},
  and find substantial benefits for direct-style.
\item
  Does our \texttt{rsyscall}-based direct-style process creation interface
  have acceptable performance overhead relative to the alternatives?
  We evaluate a number of microbenchmarks in \ref{subprocess_bench},
  and find an acceptable level of overhead.
\item
  Does direct-style process creation perform well in the ``real world''?
  We discuss our positive experience with using direct-style process creation at \twosigma in \ref{realworld}.
\end{compactitem}
\subsection{Ease of programming with direct-style}\label{ease}
\begin{table}
\input{ease.tex}
\caption{Line counts with direct-style vs fork-style}
\label{tab:ease}
\end{table}
Does direct-style process creation on Linux allow for simpler programs than the alternatives?
To answer this, we compare the direct-style examples shown in section \ref{examples}
to programs with the same behavior implemented in fork-style.

@@

Note that the cost of creating a new child process from a parent process that is already inside a namespace
(the difference between the \texttt{clone()} and \texttt{clone(NS|PID)} parent processes)
is around 2 milliseconds, comparable to the cost of running with \texttt{rsyscall}.
Note also that the overhead of creating a new process inside a new namespace
(the difference between the \texttt{clone()} and \texttt{clone(NS|PID)} child processes)
is around 15 milliseconds; subtantially greater than the overhead of \texttt{rsyscall}.
As namespaces are widely used,
this suggests our overhead is acceptable.
\subsection{Usage in the real world}\label{realworld}
At \twosigma we have used \texttt{rsyscall} and direct-style process creation
to implement an internal distributed system deployment library, written in Python,
which we'll refer to here as Toplevel.
Toplevel is in use in production,
and is extensively used as part of testing and development.
Toplevel consists of a collection of functions and modules
for each component in our system,
which start up components using direct-style process creation,
all from a single Python parent process.

@@

Several other aspects of our implementation would also be improved by new features in the Linux kernel,
as discussed in sections \ref{execve} and \ref{cloexec}.
\subsection{Portability to other Unix systems}
Other non-Linux systems
could adopt the techniques of this paper
to provide direct-style process creation.
Currently, our focus is on Linux,
but others may wish to explore porting these techniques to other operating systems.
% TODO maybe we could get rid of this paragraph?
\subsection{Large scale open source usage}
We would like to open source libraries built on top of direct-style process creation.
From our experience using such libraries at \twosigma,
we believe this would greatly ease the development of complex systems involving processes,
and would provide a substantial demonstration of the power of direct-style process creation.
\section{Conclusions}\label{conclusions}
Direct-style process creation is much less known and much less used than fork-style and spawn-style.
We have implemented direct-style process creation for Linux.
Our implementation is immediately deployable on today's Linux systems.
We have discussed various applications of processes,
and demonstrated the use of Linux direct-style process creation
to implement them.

>>>>

\begin{document}
\newcommand{\twosigma}{{[tech company]}}
\newcommand{\githuburl}{[Github url]}
\date{}
\title{\Large \bf DRAFT: Direct-style process creation on Linux}
% \author{
% {\rm Spencer Baugh}\\
% Two Sigma
% }
\maketitle
\begin{abstract}
Traditional process creation interfaces,
such as fork and spawn,
are complex to use, limited in power, and difficult to abstract over.
We develop a new process creation interface for Linux
which allows a program to create a child process in an non-running state
and initialize the new process by operating on it from the outside.
This method of process creation results in more comprehensible programs, 
has better error handling,
is comparably efficient,
and is more amenable to abstraction.
Our implementation is immediately deployable without kernel modifications on any recent Linux kernel version.
\end{abstract}

@@

We evaluate the degree to which we meet this goal by answering the following questions:
\begin{compactitem}
\item
  Does direct-style process creation on Linux allow for simpler programs than the alternatives?
  We compare implementations of similar functionality using both direct-style and fork-style in \ref{ease},
  and find substantial benefits for direct-style.
\item
  Does our \texttt{rsyscall}-based direct-style process creation interface
  have acceptable performance overhead relative to the alternatives?
  We evaluate a number of microbenchmarks in \ref{subprocess_bench},
  and find an acceptable level of overhead.
\item
  Does direct-style process creation perform well in the ``real world''?
  We discuss our positive experience with using direct-style process creation at \twosigma{} in \ref{realworld}.
\end{compactitem}
\subsection{Ease of programming with direct-style}\label{ease}
\begin{table}
\input{ease.tex}
\caption{Line counts with direct-style vs fork-style}
\label{tab:ease}
\end{table}
Does direct-style process creation on Linux allow for simpler programs than the alternatives?
To answer this, we compare the direct-style examples shown in section \ref{examples}
to programs with the same behavior implemented in fork-style.

@@

Note that the cost of creating a new child process from a parent process that is already inside a namespace
(the difference between the \texttt{clone()} and \texttt{clone(NS|PID)} parent processes)
is around 2 milliseconds, comparable to the cost of running with \texttt{rsyscall}.
Note also that the overhead of creating a new process inside a new namespace
(the difference between the \texttt{clone()} and \texttt{clone(NS|PID)} child processes)
is around 15 milliseconds; subtantially greater than the overhead of \texttt{rsyscall}.
As namespaces are widely used,
this suggests our overhead is acceptable.
\subsection{Usage in the real world}\label{realworld}
At \twosigma{} we have used \texttt{rsyscall} and direct-style process creation
to implement an internal distributed system deployment library, written in Python,
which we'll refer to here as Toplevel.
Toplevel is in use in production,
and is extensively used as part of testing and development.
Toplevel consists of a collection of functions and modules
for each component in our system,
which start up components using direct-style process creation,
all from a single Python parent process.

@@

Several other aspects of our implementation would also be improved by new features in the Linux kernel,
as discussed in sections \ref{execve} and \ref{cloexec}.
\subsection{Portability to other Unix systems}
Other non-Linux systems
could adopt the techniques of this paper
to provide direct-style process creation.
Currently, our focus is on Linux,
but others may wish to explore porting these techniques to other operating systems.
% TODO maybe we could get rid of this paragraph?
\subsection{Large scale open source usage}
We would like to open source libraries built on top of direct-style process creation.
From our experience using such libraries at \twosigma{},
we believe this would greatly ease the development of complex systems involving processes,
and would provide a substantial demonstration of the power of direct-style process creation.
\section{Conclusions}\label{conclusions}
Direct-style process creation is much less known and much less used than fork-style and spawn-style.
We have implemented direct-style process creation for Linux.
Our implementation is immediately deployable on today's Linux systems.
We have discussed various applications of processes,
and demonstrated the use of Linux direct-style process creation
to implement them.
